---
title: 'R&D in US Educational Institutions: Allocation of NSF Grants 2002 - 2015'
output: ioslides_presentation
authors: Sahar Alhassan, Kaitlyn Lynes, Deepa Mehta, Ilya Perepelitsa
widescreen: yes
---
## R&D in US Educational Institutions: Allocation of NSF Grants 2002 - 2015    

Presented by Sahar Alhassan, Kaitlyn Lynes,   
Deepa Mehta, & Ilya Perepelitsa  
December 14, 2015  

## Project Background 

* National Science Foundation (NSF) is a key engine that funds US-based R&D
* Our group focused on NSF grant dollars allocated to US institutions from 2002 to 2015
* We focused on the relationships between grant type (new, continuing, revised) over time
* We also examined the types of approved grant projects, as well as institutions and disciplines that received federal funding
* We also began to explore geographic concentration of grant funding  

## Research Method 

* Started with the NSF Grant Database 
* Collated yearly data into a separate subsets of a dataframe  
* Aggregated data by year using loops and packaged
* Developed a series of plots 
 
```{r}
load("NSF.RData") 
library(ggplot2) 
newdata <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2014-01-01'), ] 
newdata1 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2013-01-01'), ] 
newdata2 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2012-01-01'), ] 
newdata3 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2008-01-01'), ] 
newdata4 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
newdata5 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
```

## NSF Monthly Grant Allocation 2014-2015 
* One year snapshot
* Continued and New Projects Received More Grant Allocation than Revised Projects
```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
load("NSF.RData")
newdata <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2014-01-01'), ] 
newdata1 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2013-01-01'), ] 
newdata2 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2012-01-01'), ] 
newdata3 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2008-01-01'), ] 
newdata4 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
newdata5 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
ggplot(newdata, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type))
```

## NSF Grant Dollar Allocation by Month over a Two-Year Period: 2013-2015
* We decided to look at this over two years 
* This showed us a trend 
```{r, echo=FALSE, warning=FALSE}
ggplot(newdata1, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type)) 
```

## NSF Grant Dollar Allocation by Month over a Thirteen-Year Period: 1999-2015
* We looked at this trend over 16 years 
* Towards the end of every year, we see more grant dollars allocated 
```{r, echo=FALSE, warning=FALSE}
ggplot(newdata5, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type)) 
```

## Two Hypotheses
H1: Less expensive grants are more likely to be granted at the beginning of the year, while more expensive grants are more likely to be granted at the end of the year. This reflects difficulty.
H2: Pressure to spend end of fiscal year Congress Appropriations results in increased demand for grants. NSF more likely to choose expensive grants to appropriate money more quickly.

## Types of NSF Grants 2002-2015 by Action Type 
* We broke it this trend down by action type
* Revised grants saw a large gap between 2001 and 2004
```{r, echo=FALSE, warning=FALSE}
ggplot(newdata4, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type))+ facet_wrap(~action_type)
```

## Types of Research by Recipient Types 
* During this period, most grants were in the Engineering, Mathematics and Geosciences 
* Institutions appear to receive most of the grants directly 
```{r, echo=FALSE, warning=FALSE, fig.height=8, fig.width=6}
ggplot(schoolsall, aes(x=cfda_program_title, y=log(costmon), fill = cfda_program_title)) + geom_jitter(alpha=I(1/2), aes(color=cfda_program_title)) + facet_wrap(~recipient_type) +guides(colour = guide_legend(override.aes = list(size=12)))
```

## Types of Research 
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=action_type, y=log(costmon), fill = action_type)) + geom_jitter(alpha=I(1/2), aes(color=action_type)) + facet_wrap(~cfda_program_title) + guides(colour = guide_legend(override.aes = list(size=12)))
```
  
## NSF Grant Dollar Allocation by Month as a Function of Research Duration
* Are projects with a longer duration allocated larger grants? 
* Newer grants appear to be shorter in duration 
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(y=res_duration, x=log(costmon))) + geom_point(aes(color = action_type))
```

## NSF Grant Dollars by Research Duration & By Type
* Revised grants are longer than continuing grants
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(y=res_duration, x=log(costmon))) + geom_point(aes(color = action_type))+ facet_wrap(~action_type)
```

## Gaps in Research Funding: 2004
* What accounts for the gap in data in 2004?
* Is this a gap in data or was research funding halted due to a government shutdown in 2004?  
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall,aes(x=obligation_action_date, y = log(costmon))) + geom_point(aes(color= action_type )) 
```

## Density of Grant Types
* Bimodal distribution in revised grants
* New grants have higher allocations 
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6)
```

## Density of Grants - By Subject
* More density among grants for STEM fields 
* High density of grant allocation to polar programs
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=cfda_program_title)) + geom_density(alpha = 0.6) + facet_wrap(~cfda_program_title)
```

## Density of Grants - By Recipient Type
*Appears to be uniform, although some variation in Small Business recipients 
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=recipient_type)) + geom_density(alpha = 0.6) + facet_wrap(~recipient_type)
```

## Density of Grants by Action Type and Recipient Type
* Grants appear to be concentrated in certain recipient types for revised grants (i.e. State Government)
* We see narrow concentrations spiking for Small Business Grant recipients
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6) + facet_wrap(~recipient_type)
```


## Density of Grants by Action Type and by States 
* Trends appear to be consistenet in all states and territories
* Territories appear to have more new grants within this period 
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6) + facet_wrap(~principal_place_state_code)
```

## Density of Grants by Action Type and by Subject  
* New grants in relation to continuing grants  
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6) + facet_wrap(~cfda_program_title)
```

## Aggregate Grants by State
* CA, DC, MA, and NY have the most grants
* CA received 15% of the grants (consistently)
```{r, echo=FALSE, warning=FALSE}
sumcolbystate <- aggregate(fed_funding_amount ~ principal_place_state_code + fiscal_year, schoolsall, sum)
ggplot(sumcolbystate, aes(x=fiscal_year, y=fed_funding_amount)) + geom_line(aes(colour = principal_place_state_code))+ facet_wrap(~principal_place_state_code) 
```

## State Summaries by Research Type
```{r, echo=FALSE, warning=FALSE}
bigstates <- schoolsall[which(schoolsall$principal_place_state_code == "CA" | schoolsall$principal_place_state_code == "CO" | schoolsall$principal_place_state_code == "DC" | schoolsall$principal_place_state_code == "IL"| schoolsall$principal_place_state_code == "MA"| schoolsall$principal_place_state_code == "NY"| schoolsall$principal_place_state_code == "PA"| schoolsall$principal_place_state_code == "TX"), ]
ggplot(bigstates, aes(x=cfda_program_title, y=log(costmon), fill = cfda_program_title)) + geom_jitter(alpha=I(1/2), aes(color=cfda_program_title)) + facet_wrap(~principal_place_state_code) +guides(colour = guide_legend(override.aes = list(size=12)))
```

## Preliminary Conclusions and Next Steps for this Research Project
* Improving the accessibility of this data 
* Prepare the dataframe in a way that allows us to generate reports by research type, institution type, state, and key words in grant descriptions
* Be able to expand this data to other governmental datasets 

## Thank You!
