---
title: 'R&D in US Educational Institutions: Allocation of NSF Grants 2002 - 2015'
output: ioslides_presentation
authors: Sahar Alhassan, Kaitlyn Lynes, Deepa Mehta, Ilya Perepelitsa
widescreen: yes
---
## R&D in US Educational Institutions: Allocation of NSF Grants 2002 - 2015    

Presented by Sahar Alhassan, Kaitlyn Lynes,   
Deepa Mehta, & Ilya Perepelitsa  
December 14, 2015  

## 1. Project Background 

* National Science Foundation (NSF) is a key engine that funds US-based R&D
* Our group focused on NSF grant dollars allocated to US institutions from 2002 to 2015
* We focused on the relationsships between grant type (new, continuing, revised) over time
* We also examined the types of approved grant projects, as well as institutions and disciplines that received federal funding
* We also began to explore geographic concentration of grant funding  

## 2. Research Method 

* Started with the NSF Grant Database 
* Collated yearly data into a single dataset for 2002 - 2015
* Developed a series of Plots

```{r}
load("NSF.RData") 
library(ggplot2) 
newdata <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2014-01-01'), ] 
newdata1 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2013-01-01'), ] 
newdata2 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2012-01-01'), ] 
newdata3 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2008-01-01'), ] 
newdata4 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
newdata5 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
```

## 3. NSF Monthly Grant Allocation 2014-2015 
* FINDING
* FINDING
```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
load("NSF.RData")
newdata <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2014-01-01'), ] 
newdata1 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2013-01-01'), ] 
newdata2 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2012-01-01'), ] 
newdata3 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '2008-01-01'), ] 
newdata4 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
newdata5 <- schoolsall[which(schoolsall$starting_date < '2015-01-01' & schoolsall$starting_date > '1999-01-01'), ] 
ggplot(newdata, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type))
```

## 4. NSF Grant Dollar Allocation by Month over a Two-Year Period: 2013-2015
* FINDING
* FINDING
```{r, echo=FALSE, warning=FALSE}
ggplot(newdata1, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type)) 
```

## 5. NSF Grant Dollar Allocation by Month over a Thirteen-Year Period: 2002-2015
* FINDING
* FINDING
```{r, echo=FALSE, warning=FALSE}
ggplot(newdata1, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type)) 
```

## 6. Two Hypotheses
1. More research grants were allocated at year-end
2. 

## 7. Types of NSF Grants 2002-2015
```{r, echo=FALSE, warning=FALSE}
ggplot(newdata4, aes(y=log(costmon), x=starting_date)) + geom_point(aes(color = action_type))+ facet_wrap(~action_type)
```

## 8. NSF Grant Dollar Allocation by Month as a Function of Research Duration
* Are projects with a longer duration allocated larger grants? 
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(y=res_duration, x=log(costmon))) + geom_point(aes(color = action_type))
```

## 9. NSF Grant Dollars by Research Duration & By Type
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(y=res_duration, x=log(costmon))) + geom_point(aes(color = action_type))+ facet_wrap(~action_type)
```

## 10. Gaps in Research Funding: 2004
* What accounts for the gap in data in 2004?
* Is this a gap in data or was research funding halted due to a government shutdown in 2004?  
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall,aes(x=obligation_action_date, y = log(costmon))) + geom_point(aes(color= action_type )) 
```

## 11. Density of Grant Types
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6)
```

## 12. Density of Grants - By Subject
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=cfda_program_title)) + geom_density(alpha = 0.6) + facet_wrap(~cfda_program_title)
```

## 13. Density of Grants - By Recipient Type
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=recipient_type)) + geom_density(alpha = 0.6) + facet_wrap(~recipient_type)
```



## 14. Density of Grants by Action Type and Recipient Type
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6) + facet_wrap(~recipient_type)
```


## 15. Density of Grants by Action Type and by States 
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6) + facet_wrap(~principal_place_state_code)
```


## 16. Density of Grants by Action Type and by Subject  
```{r, echo=FALSE, warning=FALSE}
ggplot(schoolsall, aes(x=log(costmon), fill=action_type)) + geom_density(alpha = 0.6) + facet_wrap(~cfda_program_title)
```


## 17. OTHER???

CODE??

## 18. Discussion

* Finding
* Finding 

## 19. Preliminary Conclusions and Next Steps

* Finding
* Finding 

## 20. Thank You!
